\myTitle

### 2023-11-03 body-text-of-paper.tex
Add another GSP.  Also, some tiny edits

Genetic clustering algorithms are a commonly used approach to describe genetic structure among populations, reflecting the cumulative 
effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations 
is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker 
set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently 
interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location 
and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications, 
such as influencing evolutionary processes, maintaining genetic variation through gene flow, and driving disease dynamics 
\citep{huestis2019windborne} and affecting patterns of invasive species expansion CITE.

@@

A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
parallel analyses on simulated data where the truth is known, structured to mirror the empirical data set
as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
observed allele frequencies (see CITE, CITE, CITE).
However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed 
samples, and
inflates the perceived ability to correctly assign individuals back to source populations. \citet{anderson2008improved}
documented this type of artifact in the context of simulations to assess the power for
genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
all their ancestry from a single subpopulation.
, a type of clustering application with the limitation that individuals are fully assigned to single subpopulations (ignoring the possibility of 
mixed ancestry).
We demonstrate, below, that similar artifacts occur when simulating data to
assess how accurately admixture fractions can be estimated.  We refer to this phenomenon as resampling-induced
power inflation, or RIPI for short.
NOTE I would proposed holding on to the next sentence as we build the idea of mixed ancestry in the next paragraph END NOTE
In this study, we additionally develop and illustrate an approach by which RIPI can be eliminated
from simulations by constraining the genetic differences among simulated populations to match
those of populations from which the samples were obtained.

@@

The call the structure
we use to do this sort of cross-validation simulation of admixed individuals
a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
a GSP includes a set of founders (individuals that have no parents specified in the
pedigree), and it includes non-founders; however, we introduce an additional node
type in the GSP that represents samples taken from the non-founders. These samples
are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
any inbreeding loop indicates that more than one copy of a gene in an ancestor may
be created amongst its descendants, which is a type of sampling without
replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
samples at the sample nodes should be equal to the amount of material in the founders,
to ensure that all available data is being used to estimate power for the estimation of
admixture fractions.

@@

In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses; however, \gscramble{} is designed
to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
to keep in mind a handful of requirements for a GSP to be valid. Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\
\end{aligned}
\end{equation}
Before simulating from a GSP, \gscramble{} checks to ensure the GSP
satisfies these validity conditions.  The simulation of genomic segments
from founders to samples in a GSP is performed by the \gscramble{} function
{\footnotesize\tt segregate()}, which requires as input a list of GSPs through which
to segregate genomic segments, and, for each such GSP, a table that maps the
population labels on each GSP (in the inverted triangles atop the founders) to the
different sampled groups of individuals.

>>>>

Genetic clustering algorithms are commonly used to describe genetic structure among populations, reflecting the cumulative 
effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations 
is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker 
set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently 
interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location 
and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications, 
such as influencing evolutionary processes, maintaining genetic variation through gene flow, and driving disease dynamics 
\citep{huestis2019windborne} and affecting patterns of invasive species expansion CITE.

@@

A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to conduct
parallel analyses on simulated data where the truth is known, structured to mirror the empirical data set
as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.
Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
observed allele frequencies (see CITE, CITE, CITE).
However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed 
samples, and
inflates the perceived ability to correctly assign individuals back to source populations. \citet{anderson2008improved}
documented this type of artifact in the context of simulations to assess the power for
genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
all their ancestry from a single subpopulation.
We demonstrate, below, that similar artifacts occur when simulating data to
assess how accurately admixture fractions can be estimated.  We refer to this phenomenon as resampling-induced
power inflation, or RIPI for short.
NOTE I would proposed holding on to the next sentence as we build the idea of mixed ancestry in the next paragraph END NOTE
In this study, we additionally develop and illustrate an approach by which RIPI can be eliminated
from simulations by constraining the genetic differences among simulated populations to match
those of populations from which the samples were obtained.

@@

We call the structure
we use to do this sort of cross-validation simulation of admixed individuals
a {\em genomic simulation pedigree} or GSP.  Like any pedigree,
a GSP includes a set of founders (individuals that have no parents specified in the
pedigree), and it includes non-founders; however, we introduce an additional node
type in the GSP that represents samples taken from the non-founders. These samples
are the simulated admixed individuals.  A GSP must have no inbreeding loops, since
any inbreeding loop indicates that more than one copy of a gene in an ancestor may
be created amongst its descendants, which is a type of sampling without
replacement.  Furthermore, in a GSP, the amount of genetic material taken from the
samples at the sample nodes should be equal to the amount of material in the founders,
to ensure that all available data is being used to estimate power for the estimation of
admixture fractions.

@@

In the \gscramble{} package there is a convenience function called {\footnotesize\tt create\_GSP()} that
will create GSPs for samples up to and including second-generation backcrosses (an example appears in Figure~\ref{fig:exampleGSPs}); 
%%%%%
\begin{figure}
\newcommand{\gspcaptwo}{\footnotesize
An example GSP generated by {\footnotesize\tt create\_GSP()} that produces four different
admixture categories: one~$F_1$ in sample node~s7; two~$F_2$'s in sample node~s11; one $BC_1$ in sample node~s9; and two $BC_2$'s in sample node~s10. 
}
\begin{center}
\includegraphics[width=\columnwidth]{figures/AllThem.pdf}
\end{center}
\caption[\gspcaptwo]{\gspcaptwo}
\label{fig:gsp1}
\end{figure}
%%%%
however, \gscramble{} is designed
to handle user-specified GSPs of arbitrary complexity.  When creating one's own GSP, it is useful
to keep in mind a handful of requirements for a GSP to be valid. Writing $\founders$ for the set of all founder nodes, $\nonfound$ for the set of all non-founder nodes, and $\nonfounds$ for all non-founder
nodes that are not parents of other non-founders, but are adjacent to a sample node, $\nonfounde$ for all
non-founder nodes that have non-founder node daughters, but no sample nodes, and $\nonfoundse$ for all
non-founder nodes that are parents of non-founder nodes and also adjacent to sample-nodes, the necessary conditions
for a valid GSP, in addition to having no inbreeding loops, are:
\begin{equation}
\begin{aligned}
\sum_{x \in \nonfounds \cup \nonfoundse} S_x &= |\founders| &  \\
\sum_{i=1}^{e_x} G^-_{x,i} &= 2 & \forall x \in \founders \\
G^+_{x,1} &= G^+_{x,2} & \forall x \in \nonfound \\
\sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & \forall x \in \nonfounde  \\
 S_x/2 + \sum_{i=1}^{e_x} G^-_{x,i} &= G^+_{x,1} +  G^+_{x,2} & ~~~\forall x \in \nonfoundse.  \\
\end{aligned}
\end{equation}
Before simulating from a GSP, \gscramble{} checks to ensure the GSP
satisfies these validity conditions.  The simulation of genomic segments
from founders to samples in a GSP is performed by the \gscramble{} function
{\footnotesize\tt segregate()}, which requires as input a list of GSPs through which
to segregate genomic segments, and, for each such GSP, a table that maps the
population labels on each GSP (in the inverted triangles atop the founders) to the
different sampled groups of individuals.
